Automatic Singing Transcription: Phonemes and Alignment

### 2023-12-05 report/paper.tex
minor organizational paper changes

\title{Automatic Singing Transcription}
\author{Victor Shan}

@@

\subsection{Dataset Augmentation}
\subsubsection{SpecAugment}
Due to the lack of datasets, it is more important to make the most of the existing datasets. One
way to do this is to augment the existing datasets. SpecAugment is a series of techniques that
augment the audio spectrogram to improve the performance of ASR models. A frequent intermediate
representation are Mel-Frequency Cepstral Coefficients (MFCC).

@@

% ----------------- INCLUDE MFCC IMAGE -----------------

@@

\subsubsection{Transforming Utterance level datasets into Phoneme level datasets} \label{sec:utteranceToPhoneme}
A technique from ASR that can also be applied to AST is transforming utterance/sentence
level datasets into phoneme level datasets. There are many ASR datasets that contain single utterances
such as LibriSpeech and MUSDB18 \citep{Librispeech,musdb18}. These datasets can be effectively
transformed into phoneme level datasets by using a phoneme dictionary such as the CMU Pronouncing
Dictionary \citep{CMUDict}. This dictionary contains a mapping from words to phonemes. Using this
dictionary, the utterances can be transformed into phoneme sequences. These sequences can then be
used with a Connectionist Temporal Classification (CTC) loss function to train AST models \citep{CTC}.
\begin{figure*}
    \centering
    \includegraphics[width=0.5\textwidth]{assets/CTC.png}
    \caption{CTC Function collapsing repetitions \citep{CTC}}
    \label{fig:CTC}
\end{figure*}
CTC allows for the model to output a sequence of phonemes per time step and the
repetitions are collapsed into a single phoneme as shown in Figure \ref{fig:CTC}.
This technique was used to allow a model to incorperate the utterance level LibriSpeech
dataset into the training for a phoneme ASR model \citep{wav2vec}. Timestamps can be
retrieved from the pre-CTC output that had time-aligned phoneme classifications.
The same process can be applied to AST models to allow them to use the utterance level song datasets.

@@

\subsubsection{Children's Songs Dataset}
This is a dataset of 50 English and 50 Korean children's songs with grapheme level timestamps and
pitches. This dataset was sung by a single professional Korean singer in two separate keys for each
song \citep{ChildrensSongs}. This dataset is a good candidate for training AST models because it
includes the onset and offset information, pitch, and grapheme level timestamps that could be
preprocessed to phoneme or word level timestamps. It is also a very clean dataset because of the
professional performance and setting. The downside of this dataset is that graphemes are
not as popular as phonemes, word or utterance labels in either ASR or AST.

@@

\subsection{Preprocessing Datasets}
To preprocess datasets, the audio files were preprocessed in the same way WhisperX preprocessed
data with the VAD Cut \& Merge technique \citep{whisperX} except with a slight modification before
trying to detect voices. Since the audio files were songs, most of them will have both singing and
background music. This means that the VAD model will have a harder time detecting when the singing
stops. To help the VAD model, the audio files were first passed through Demucs v4 \citep{Demucs}
to separate the vocals from the background music. Then, the VAD model was used to detect the
voices. Finally, the audio files were cut into segments at the detected voice sections. To reduce
memory requirements and because lyrics utterances are often shorter than 30 seconds, the segments
will be cut into 8-15 second segments. Segments that are shorter than 8 seconds can either be padded
or merged with other segments into segments of 8-15 seconds.

@@

\subsection{Augmenting Datasets}
There are many ways to augment datasets. The three ways mentioned in SpecAugment includes
time warping, frequency masking and time masking \citep{SpecAugment}. These techniques can be
applied to the audio spectrogram before it is fed into the model. Time warping is a technique that
stretches or compresses the audio spectrogram in the time dimension so it must be handled carefully
to update the labels accodingly. Another way to augment the datasets is to transform the audio
pitch. This can be done by shifting the pitch up or down.

@@

\subsection{VAD Performance}

@@

\subsection{Lyric Generation Performance}
% WER

@@

\section{Future Work}
\subsection{Adversarial Training}
\subsection{Transform Singing to Speech Data}
remove pitch from singing
\subsection{Phoneme to Grapheme}
\citep{G2P} or \citep{P2G}

>>>>

\title{Automatic Singing Transcription: Phonemes and Alignment}
\author{Victor Shan}

@@

\subsection{Dataset Augmentation}
\subsubsection{SpecAugment}
Due to the lack of datasets, it is more important to make the most of the existing datasets. One
way to do this is to augment the existing datasets. SpecAugment is a series of techniques that
augment the audio spectrogram to improve the performance of ASR models. A common variation of this
representation is the Mel-Frequency Cepstral Coefficient (MFCC) spectrogram.

@@

{
    \centering
    \includegraphics[width=0.45\textwidth]{assets/specaugment.png}
}

@@

\subsubsection{Transforming Utterance level datasets into Phoneme level datasets} \label{sec:utteranceToPhoneme}
A technique from ASR that can also be applied to AST is transforming utterance/sentence
level datasets into phoneme level datasets. There are many ASR datasets that contain single utterances
such as LibriSpeech and JamendoLyrics. These datasets can be effectively
transformed into phoneme level datasets by using a phoneme dictionary such as the CMU Pronouncing
Dictionary \citep{CMUDict}. This dictionary contains a mapping from words to phonemes. Using this
dictionary, the utterances can be transformed into phoneme sequences. These sequences can then be
used with a Connectionist Temporal Classification (CTC) loss function to train AST models \citep{CTC}.
\begin{figure*}
    \centering
    \includegraphics[width=0.5\textwidth]{assets/CTC.png}
    \caption{Example of CTC Function collapsing repetitions. Image from \citep{CTC}}
    \label{fig:CTC}
\end{figure*}
CTC allows for the model to output a sequence of phonemes per time step and the
repetitions are collapsed into a single phoneme as shown in Figure \ref{fig:CTC}.
This technique was used to allow a model to incorperate the utterance level LibriSpeech
dataset into the training for a phoneme ASR model \citep{wav2vec}. Timestamps can be
retrieved from the pre-CTC output that had time-aligned phoneme classifications.
The same process can be applied to AST models to allow them to use the utterance level song datasets.

@@

\subsubsection{Children's Songs Dataset}
This is a dataset of 50 English and 50 Korean children's songs with grapheme level timestamps and
pitches. This dataset was sung by a single professional Korean singer in two separate keys for each
song \citep{CSD}. This dataset is a good candidate for training AST models because it
includes the onset and offset information, pitch, and grapheme level timestamps that could be
preprocessed to phoneme or word level timestamps. It is also a very clean dataset because of the
professional performance and setting. The downside of this dataset is that graphemes are
not as popular as phonemes, word or utterance labels in either ASR or AST.

@@

\subsection{Preprocessing Datasets}
To preprocess datasets, the audio files were preprocessed in the same way WhisperX preprocessed
data with the VAD Cut \& Merge technique \citep{whisperX} except with a slight modification before
trying to detect voices. Since the audio files were songs, most of them will have both singing and
background music. This means that the VAD model will have a harder time detecting when the singing
stops. To help the VAD model, the audio files were first passed through Demucs v4 \citep{Demucs}
to separate the vocals from the background music. Then, the VAD model was used to detect the
voices. Finally, the audio files were cut into segments at the detected voice sections. To reduce
memory requirements and because lyrics utterances are often shorter than 30 seconds, the segments
will be cut into 5-15 second segments. Smaller segments can either be padded
or merged with other segments into larger ones.

@@

\subsection{Augmenting Datasets}
There are many ways to augment datasets. The three ways mentioned in SpecAugment includes
time warping, frequency masking and time masking \citep{SpecAugment}. These techniques can be
applied to the audio spectrogram before it is fed into the model. Time warping is a technique that
stretches or compresses the audio spectrogram in the time dimension due to the complexity of
realigning the audio. We use time and frequency maksing for training our models.

@@

% \subsection{VAD Performance}

@@

% \subsection{Lyric Generation Performance}
% % WER

@@

\section{Future Work}
\subsection{Train on More Datasets}
\subsection{Finetuning Wav2Vec 2.0 Feature Extractor}
\subsection{Adversarial Training}
\subsection{Transform Singing to Speech Data}
remove pitch from singing
\subsection{Phoneme to Grapheme}
\citep{G2P} or \citep{P2G}
